### 2020-05-22 ms/suppmat.tex
Add Supplementary Materials

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\onehalfspacing

\usepackage[round]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\usepackage[utf8]{inputenc}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark 
\usepackage{rotating}
%\usepackage[nolists,tablesfirst]{endfloat}
%\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables

% omit bibliography (\nobibliography{...})
\usepackage{bibentry}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage[detect-all]{siunitx} % for SI units

\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
\setlength{\LTcapwidth}{\textwidth}

% custom hyphenation:
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_r$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
% Select what to do with command \comment:  
\newcommand{\comment}[1]{}  % comment not shown
%\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

%\bibliographystyle{model2-names}
\bibliographystyle{fishfishnourl}

%%% MY MACROS

% Make supplementary tables and figures start with S1 and T2
% use \beginsupplement to reset count
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\begin{document}
\linenumbers

\noindent
{\Large Supplementary Materials}
\newline

\noindent
{\Large for}
\newline

\noindent
{\large Trends in marine survival of Atlantic salmon populations in eastern Canada}
\newline

\noindent
Sebasti\'{a}n A. Pardo\textsuperscript{1*}, Geir H. Bolstad\textsuperscript{2}, J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1}
\newline

\noindent
\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
\beginsupplement

This Supplementary Material includes additional tables to accompany the publication.

\clearpage

\subsection*{Raw smolt data used in the model}

\input{smolts_tab.tex}

\input{smoltscv_tab.tex}

\subsection*{Annual smolt abundances estimates derived from the Bayesian hierarchical model}

%\begin{table}[ht]
%\centering
%\begin{longtable}{lrrrrrr}
%\caption{Posterior annual smolt estimates for the seven populations examined. These
%                    estimates differ from the smolt data used as they include both observation error and shrinkage from
%                    the hierarchical model specification.} 
%\hline
\input{tableS1_smolts}
%\end{longtable}
%\end{table}
\clearpage

\subsection*{Population-level estimates of proportion returning after one winter at sea}

\input{tableS2_smu}

\hspace{8em}

\input{tableS3_lsmu}

\input{tableS4_lssig}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.9\linewidth]{simul_timeseries_declining.pdf} \caption{Simulated
%         time series of returning adult salmon abundance in the six scenarios. The lines denote the
%     simulated abundance estimates without observation error while the points are the same estimates
% including observation error.}
% \label{fig:simul-ts-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.95\linewidth]{sir_out_18_declining.pdf} \caption{Yearly
%         estimated \So, \St, and \Pg values in the six scenarios. True values
%         are denoted by blue circles, black circles show median estimates,
%         error bars indicate the 25\% and 75\% quantiles.}
% \label{fig:sir-out-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.7\linewidth]{simul_plots_6_declining.pdf}
%     \caption{Comparison of estimated and true \So values in the six 
%         scenarios. True \So values are deterministic, black circles show
%         median \So estimates, error bars indicate the  25\% and 75\%
%         quantiles, while the blue line denotes a linear model fit of the
%         medians. The one-to-one relationship is shown by the gray dashed
%         line.} 
% \label{fig:simul-six-plots-dec} 
% \end{figure}

%\clearpage
%\subsection*{References}
\nobibliography{subset}

\end{document}