### 2021-04-04 ms/suppmat.tex
Add changes up to ICES JMS submission

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\onehalfspacing

\usepackage[round]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\usepackage[utf8]{inputenc}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark 
\usepackage{rotating}
%\usepackage[nolists,tablesfirst]{endfloat}
%\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables

% omit bibliography (\nobibliography{...})
\usepackage{bibentry}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage[detect-all]{siunitx} % for SI units

\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
\setlength{\LTcapwidth}{\textwidth}

% custom hyphenation:
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_r$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
% Select what to do with command \comment:  
\newcommand{\comment}[1]{}  % comment not shown
%\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

%\bibliographystyle{model2-names}
\bibliographystyle{fishfishnourl}

%%% MY MACROS

% Make supplementary tables and figures start with S1 and T2
% use \beginsupplement to reset count
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\begin{document}
\linenumbers

\noindent
{\Large Supplementary Materials}
\newline

\noindent
{\Large for}
\newline

\noindent
{\large Trends in marine survival of Atlantic salmon populations in eastern Canada}
\newline

\noindent
Sebasti\'{a}n A. Pardo\textsuperscript{1*}, Geir H. Bolstad\textsuperscript{2}, J. Brian Dempson\textsuperscript{3}, 
        Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3}, 
        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1}
\newline

\noindent
\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
\beginsupplement

This Supplementary Material includes additional tables to accompany the publication.

\clearpage

\subsection*{Estimation of uncertainty in return estimates based on scale data}

We incorporated the uncertainty in the conversion from size group (small and large returns) to age class (1SW and 2SW returns) by 
approximating the variance of log return abundances of each maiden age class (i.e. 1SW and 2SW) assuming that the probability 
of the true number of returns of a given age, river, year, and size group ($N_{r,t,s,a}$) can be derived by bootstrapping a distribution
of number of returns from the hypergeometric distribution

% \begin{equation}
% p[N_{r,t,s,a}|0:N_{r,t,s}] \sim Hypergeometric(n_{r,t,s}, 0:N_{r,t,s}, N_{r,t,s}:0, n_{r,t,s,a}),
% \end{equation}

\begin{equation}
P(N_{obs,r,t,s,a}|n_{r,t,s,a}) \sim Hypergeometric(N_{r,t,s,a}, N_{r,t,s} - N_{r,t,s,a}, n_{r,t,s}),
\end{equation}

where $P(N_{obs,r,t,s,a}|n_{r,t,s,a})$ is the probability of having a specific number of returns $N_{r,t,s,a}$ given 
$n_{r,t,s,a}$, which is the number of scale samples in river r, year t, of returning age a, and of size group s;
$N_{r,t,s}$ is the total number of returns in river r, year t, and of size group s;
$N_{r,t,s} - N_{r,t,s,a}$ is the total numbers of returns in river r, year t, of size group s that are \textbf{not} of returning age a;
and $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s.
We iterated through all possible values of $N_{r,t,s,a}$ (i.e. from zero to $N_{r,t,s}$), to obtain a probavility for each value, which
we then used as weighting probabilities to sample 10,000 random draws of $N_{r,t,s,a}$, which we denote as $\hat{N}_{r,t,s,a}$. 
We then add the draws of the same age from both size groups to obtain a vector of estimates of number of returns of a given age a (i.e. 1SW or 2SW) for a given river and year ($\hat{N}_{r,t,a}$):

% k is the number of successes ($n_{r,t,s,a}$) (successes)
% N is the total number of individuals of size s ($N_{r,t,s}$) (total count) 
% K is the number of age a individual of size s ($N_{r,t,s,a}$) (total successes)
% n is the total number of scale samples ($n_{r,t,s}$) (trials)

% where $n_{r,t,s,a}$ is the number of scale samples in river r, year t, of returning age a, and of size group s;
% $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s; and
% $N_{r,t,s}$ is the total number of returns in river, year t, and of size group s.

% We therefore obtain a vector of the probabilities of the returns of a given river r, year t, size group s and age a (i.e. $N_{r,t,s,a}$)
% across the whole range of potential values, i.e. from zero to $N_{r,t,s}$, which provides the weighting probabilities to generate 10,000 random draws of $N_{r,t,s,a}$.

\begin{equation}
    \hat{N}_{r,t,a} = \hat{N}_{r,t,a,s=small }+ \hat{N}_{r,t,a,s=large}.
\end{equation}

Lastly, we obtain the estimates of $\epsilon_{r,t,a}$ by calculating the standard deviation of log-transformed $\hat{N}_{r,t,a}$:

\begin{equation}
\epsilon_{r,t,a} = \sigma_{log(\hat{N}_{r,t,a})}.
\end{equation}

The $\epsilon_{r,t,a}$ parameter is used as the variance term in equations~8 and 9 of the manuscript.
The data used is in Table~\ref{tab:scales} while the estimates of $\epsilon_{r,t,a}$ are provided in Tables~\ref{tab:epsilon1} and~\ref{tab:epsilon2}.

There are a number of years with very few scales samples, particularly for the
Campbellton population, and one year for this population (2011) with no scales
samples. Under these conditions, the model provides high estimates of
$\epsilon$, but which are somewhat constrained by the log-transformation of
abundance draws of a uniform distribution. Thus, the value of $\epsilon$ for
2011, the year with no scale samples, is of 0.711 (Table~\ref{tab:epsilon1}),
which is likely an underestimate of true uncertainty given the absence of
data. Nonetheless, the model results in small $\epsilon$ values for years
where there are ten or more scale samples. 
For years where there are no scale nor return samples to draw inferences from (e.g. a year for which there are no small 2SW scales, nor large scale samples),
sometimes the error for 2SW returns cannot be estimated; in these cases we set the value of $\epsilon_{2}$ to 0.999.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/camp-scale-epsilon-example.png} \caption{Example
        of the estimation of $\epsilon_{1}$ for two different years of data for the Campbellton population: year 2011 where
        there were no scale samples, and year 2014 with a large number of scales samples. a) Distribution of draws of potential
    1SW abundances, adding the estimates from both small and large returns, for the year 2011. Note that the distribution spans the whole range of
returns (up to 5443). b) The same distribution but log-transformed, which results in a right-skewed distribution. The blue line represents the normal
distribution fitted to the log-transformed data, from which $\epsilon_{1}$ is equal to its standard deviation. b) Distribution of draws of potential
    1SW abundances, adding the estimates from both small and large returns, for the year 2014, which is much tightly constrained. d) The distribution
of log-transformed abundance draws for year 2014. In this case, the normal distribution fitted is much tighter and more closely resembles the underlying histogram,
and results in a much lower estimate of $\epsilon_{1}$.}
    \label{fig:priors} 
\end{figure}

\clearpage

\input{scales_tab.tex}

\clearpage

\input{epsilon1_tab.tex}

\clearpage

\input{epsilon2_tab.tex}

\clearpage

\subsection*{Raw smolt data used in the model}

\input{smolts_tab.tex}

\clearpage

\subsection*{Error in smolt abundances used in the model}

\input{smoltscv_tab.tex}

\clearpage

\subsection*{Annual smolt abundances estimates derived from the Bayesian hierarchical model}

%\begin{table}[ht]
%\centering
%\begin{longtable}{lrrrrrr}
%\caption{Posterior annual smolt estimates for the seven populations examined. These
%                    estimates differ from the smolt data used as they include both observation error and shrinkage from
%                    the hierarchical model specification.} 
%\hline
\input{tableS1_smolts}
%\end{longtable}
%\end{table}
\clearpage

\subsection*{Population-level estimates of smolt abundance}

\input{tableS2_smu}

%\hspace{8em}

\clearpage

\input{tableS3_lsmu}

\input{tableS4_lssig}

\clearpage

\subsection*{Population-level estimates of proportion returning after one winter at sea}

\input{table-logitprmu.tex}

\hspace{8em}

\input{table-logitprsigma.tex}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.9\linewidth]{simul_timeseries_declining.pdf} \caption{Simulated
%         time series of returning adult salmon abundance in the six scenarios. The lines denote the
%     simulated abundance estimates without observation error while the points are the same estimates
% including observation error.}
% \label{fig:simul-ts-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.95\linewidth]{sir_out_18_declining.pdf} \caption{Yearly
%         estimated \So, \St, and \Pg values in the six scenarios. True values
%         are denoted by blue circles, black circles show median estimates,
%         error bars indicate the 25\% and 75\% quantiles.}
% \label{fig:sir-out-dec} 
% \end{figure}

% \begin{figure}[htbp] \centering
%     \includegraphics[width=0.7\linewidth]{simul_plots_6_declining.pdf}
%     \caption{Comparison of estimated and true \So values in the six 
%         scenarios. True \So values are deterministic, black circles show
%         median \So estimates, error bars indicate the  25\% and 75\%
%         quantiles, while the blue line denotes a linear model fit of the
%         medians. The one-to-one relationship is shown by the gray dashed
%         line.} 
% \label{fig:simul-six-plots-dec} 
% \end{figure}

%\clearpage
%\subsection*{References}
\nobibliography{subset}

\end{document}