### 2021-05-24 ms/suppmat.tex
Add prior sensitivity tests

\documentclass[12pt]{article}
\usepackage[top=0.85in,left=1.0in,right=1.0in,footskip=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{setspace}
\usepackage{lineno}
\usepackage[hidelinks]{hyperref}
\onehalfspacing

\usepackage[round]{natbib}
\setcitestyle{authoryear}
\bibpunct{(}{)}{;}{a}{}{;}

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\usepackage[utf8]{inputenc}

\usepackage{xspace}
\usepackage{xfrac} % for diagonal inline fractions in text
%\usepackage{array} % for making whole row bold in table
\usepackage{colortbl} % for background colours in table rows
\usepackage{longtable}
\usepackage{amssymb} % for \checkmark
\usepackage{rotating}
%\usepackage[nolists,tablesfirst]{endfloat}
%\DeclareDelayedFloatFlavor{sidewaystable}{table}
\usepackage{makecell} % for \makecell in tables

% omit bibliography (\nobibliography{...})
\usepackage{bibentry}

\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array} % for table wrapping of columns
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage[detect-all]{siunitx} % for SI units

\makeatletter\let\expandableinput\@@input\makeatother % expandable input for \input inside tables
\setlength{\LTcapwidth}{\textwidth}

% custom hyphenation:
\hyphenation{elasmo-branchs}

% Macros
\newcommand{\So}{$S_{1}$\xspace}
\newcommand{\St}{$S_{2}$\xspace}
\newcommand{\Pg}{$P_g$\xspace}
\newcommand{\prmu}{$\mu_{Pg,r}$\xspace}
\newcommand{\Linf}{$L_{\infty}$}
\newcommand{\DWinf}{$DW_{\infty}$}
\newcommand{\alphat}{$\tilde{\alpha}$}
\newcommand{\lamat}{$l_{\alpha_{mat}}$}
\newcommand{\lamatb}{$l_{\alpha_{mat}}b$}
\newcommand{\rmax}{$r_{max}$\xspace}
\newcommand{\ageratio}{$\alpha_{mat}/\alpha_{max}$}
\newcommand{\yr}{year\textsuperscript{-1}}
\newcommand{\rsq}{$R^2$\xspace}
% Select what to do with command \comment:
\newcommand{\comment}[1]{}  % comment not shown
%\newcommand{\comment}[1]{\par {\bfseries \color{blue} #1 \par}} % comment shown
%% END MACROS SECTION

%\bibliographystyle{model2-names}
\bibliographystyle{fishfishnourl}

%%% MY MACROS

% Make supplementary tables and figures start with S1 and T2
% use \beginsupplement to reset count
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\begin{document}
\linenumbers

\noindent
{\Large Supplementary Materials}
\newline

\noindent
{\Large for}
\newline

\noindent
{\large Trends in marine survival of Atlantic salmon populations in eastern Canada}
\newline

\noindent
Sebasti\'{a}n A. Pardo\textsuperscript{1*}, Geir H. Bolstad\textsuperscript{2}, J. Brian Dempson\textsuperscript{3},
        Julien April\textsuperscript{4}, Ross A. Jones\textsuperscript{5}, %Martha J. Robertson\textsuperscript{3},
        Dustin Raab\textsuperscript{6}, Jeffrey A. Hutchings\textsuperscript{1,7,8}
\newline

\noindent
\small{\textsuperscript{1} Department of Biology, Dalhousie University, Halifax, NS, Canada\\}
\small{\textsuperscript{2} Norwegian Institute for Nature Research (NINA), Trondheim, Norway\\}
\small{\textsuperscript{3} Fisheries and Oceans Canada, St. John's, NL, Canada\\}
\small{\textsuperscript{4} Minist\`{e}re des For\^{e}ts, de la Faune et des Parcs, Qu\'{e}bec, QC, Canada\\}
\small{\textsuperscript{5} Fisheries and Oceans Canada, Moncton, NB, Canada\\}
\small{\textsuperscript{6} Fisheries and Oceans Canada, Dartmouth, NS, Canada\\}
\small{\textsuperscript{7} Institute of Marine Research, Flødevigen Marine Research Station, N-4817 His, Norway\\}
\small{\textsuperscript{8} Centre for Coastal Research, University of Agder, N-4604 Kristiansand, Norway\\}
\small{\textsuperscript{*} Corresponding author: spardo@dal.ca}
\beginsupplement

This Supplementary Material includes additional tables to accompany the publication.

\clearpage

\subsection*{Estimation of uncertainty in return estimates based on scale data}

We incorporated the uncertainty in the conversion from size group (small and large returns) to age class (1SW and 2SW returns) by
approximating the variance of log return abundances of each maiden age class (i.e. 1SW and 2SW) assuming that the probability
of the true number of returns of a given age, river, year, and size group ($N_{r,t,s,a}$) can be derived by bootstrapping a distribution
of number of returns from the hypergeometric distribution

% \begin{equation}
% p[N_{r,t,s,a}|0:N_{r,t,s}] \sim Hypergeometric(n_{r,t,s}, 0:N_{r,t,s}, N_{r,t,s}:0, n_{r,t,s,a}),
% \end{equation}

\begin{equation}
P(N_{obs,r,t,s,a}|n_{r,t,s,a}) \sim Hypergeometric(N_{r,t,s,a}, N_{r,t,s} - N_{r,t,s,a}, n_{r,t,s}),
\end{equation}

where $P(N_{obs,r,t,s,a}|n_{r,t,s,a})$ is the probability of having a specific number of returns $N_{r,t,s,a}$ given
$n_{r,t,s,a}$, which is the number of scale samples in river r, year t, of returning age a, and of size group s;
$N_{r,t,s}$ is the total number of returns in river r, year t, and of size group s;
$N_{r,t,s} - N_{r,t,s,a}$ is the total numbers of returns in river r, year t, of size group s that are \textbf{not} of returning age a;
and $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s.
We iterated through all possible values of $N_{r,t,s,a}$ (i.e. from zero to $N_{r,t,s}$), to obtain a probability for each value, which
we then used as weighting probabilities to sample 10,000 random draws of $N_{r,t,s,a}$, which we denote as $\hat{N}_{r,t,s,a}$.
We then add the draws of the same age from both size groups to obtain a vector of estimates of number of returns of a given age a (i.e. 1SW or 2SW) for a given river and year ($\hat{N}_{r,t,a}$):

% k is the number of successes ($n_{r,t,s,a}$) (successes)
% N is the total number of individuals of size s ($N_{r,t,s}$) (total count)
% K is the number of age a individual of size s ($N_{r,t,s,a}$) (total successes)
% n is the total number of scale samples ($n_{r,t,s}$) (trials)

% where $n_{r,t,s,a}$ is the number of scale samples in river r, year t, of returning age a, and of size group s;
% $n_{r,t,s}$ is the total number of scale samples in river r, year t, and of size group s; and
% $N_{r,t,s}$ is the total number of returns in river, year t, and of size group s.

% We therefore obtain a vector of the probabilities of the returns of a given river r, year t, size group s and age a (i.e. $N_{r,t,s,a}$)
% across the whole range of potential values, i.e. from zero to $N_{r,t,s}$, which provides the weighting probabilities to generate 10,000 random draws of $N_{r,t,s,a}$.

\begin{equation}
    \hat{N}_{r,t,a} = \hat{N}_{r,t,a,s=small }+ \hat{N}_{r,t,a,s=large}.
\end{equation}

Lastly, we obtain the estimates of $\epsilon_{r,t,a}$ by calculating the standard deviation of log-transformed $\hat{N}_{r,t,a}$:

\begin{equation}
\epsilon_{r,t,a} = \sigma_{log(\hat{N}_{r,t,a})}.
\end{equation}

The $\epsilon_{r,t,a}$ parameter is used as the variance term in equations~8 and 9 of the manuscript.
The data used is in Table~\ref{tab:scales} while the estimates of $\epsilon_{r,t,a}$ are provided in Tables~\ref{tab:epsilon1} and~\ref{tab:epsilon2}.

There are a number of years with very few scales samples, particularly for the
Campbellton population, and one year for this population (2011) with no scales
samples. Under these conditions, the model provides high estimates of
$\epsilon$, but which are somewhat constrained by the log-transformation of
abundance draws of a uniform distribution. Thus, the value of $\epsilon$ for
2011, the year with no scale samples, is of 0.719 (Table~\ref{tab:epsilon1}),
which is likely an underestimate of true uncertainty given the absence of
data. Nonetheless, the model results in small $\epsilon$ values for years
where there are ten or more scale samples.
For years where there are no scale nor return samples to draw inferences from (e.g. a year for which there are no small 2SW scales, nor large scale samples),
sometimes the error for 2SW returns cannot be estimated; in these cases we set the value of $\epsilon_{2}$ to 0.999.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/camp-scale-epsilon-example.png} \caption{Example
        of the estimation of $\epsilon_{1}$ for two different years of data for the Campbellton population: year 2011 where
        there were no scale samples, and year 2014 with a large number of scales samples. a) Distribution of draws of potential
    1SW abundances, adding the estimates from both small and large returns, for the year 2011. Note that the distribution spans the whole range of
returns (up to 5443). b) The same distribution but log-transformed, which results in a right-skewed distribution. The blue line represents the normal
distribution fitted to the log-transformed data, from which $\epsilon_{1}$ is equal to its standard deviation. c) Distribution of draws of potential
    1SW abundances, adding the estimates from both small and large returns, for the year 2014, which is much more tightly constrained. d) The distribution
of log-transformed abundance draws for year 2014. In this case, the normal distribution fitted is much tighter and more closely resembles the underlying histogram,
and results in a much lower estimate of $\epsilon_{1}$.}
    \label{fig:priors}
\end{figure}

\clearpage

\input{scales_tab.tex}

\clearpage

\input{epsilon1_tab.tex}

\clearpage

\input{epsilon2_tab.tex}

\clearpage

\subsection*{Raw smolt data used in the model}

\input{smolts_tab.tex}

\clearpage

\subsection*{Uncertainty in smolt abundances used in the model}

\input{smoltscv_tab.tex}

\clearpage

\subsection*{Annual smolt abundances estimates derived from the Bayesian hierarchical model}

%\begin{table}[ht]
%\centering
%\begin{longtable}{lrrrrrr}
%\caption{Posterior annual smolt estimates for the seven populations examined. These
%                    estimates differ from the smolt data used as they include both observation error and shrinkage from
%                    the hierarchical model specification.}
%\hline
\input{tableS1_smolts}
%\end{longtable}
%\end{table}
\clearpage

\subsection*{Population-level estimates of smolt abundance}

\input{tableS2_smu}

%\hspace{8em}

\clearpage

\input{tableS3_lsmu}

\input{tableS4_lssig}

\clearpage

\subsection*{Population-level estimates of proportion returning after one winter at sea}

\input{table-logitprmu.tex}

\hspace{8em}

\input{table-logitprsigma.tex}

\clearpage

\subsection*{Posterior estimates with weaker priors}

To assess the sensitivity of our model parameter estimates to the choice of
priors, we assessed posterior estimate under two weaker prior scenarios:
one with wider priors for \So, \St (Fig.~\ref{fig:zweak}), and a second one with wide priors
of \prmu for all populations regardless of whether they were considered
1SW-dominated or not (Fig.~\ref{fig:prweak}).

\begin{figure}[htbp] \centering
    \includegraphics[width=0.55\linewidth]{figures/z-weak-prior.png}
    \caption{Priors used for marine survival of one sea-winter (\So) and two
        sea-winter (\St) returns. Solid lines denote strong priors used in the
        study, while dotted lines denote weaker priors used to assess
        sensitivity to prior choice.} \label{fig:zweak} \end{figure}

\begin{figure}[htbp] \centering
    \includegraphics[width=0.55\linewidth]{figures/pr-weak-prior.png}
    \caption{Prior of the probability of returning as 1SW for all populations
        under the weak \prmu prior scenario. Note this prior is identical to
        that used in the study for non 1SW-dominated populations.}
    \label{fig:prweak} \end{figure}

The model runs with weaker priors did not differ mayorly from those with
stronger priors, particularly for posterior estimates of \So, which show the
same overall trends for all populations and some yearly \So estimates with
high uncertainty (i.e., large error bars) show greater uncertainty uncertainty
under the weak \So and \St scenario (Fig.~\ref{fig:s1-trends-priors}).
Not unexpectedly, the weak \Pg scenario had no visually discernible effect on
posterior estimates of \So.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s1-trends-priors.png}
    \caption{Median posterior estimates of survival in the first year at sea
        (\So) for the seven populations examined under three scenarios with
        different priors. The strong prior scenario is the one presented in
        our study, the weak \Pg priors uses the same weak prior for all
        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
    \label{fig:s1-trends-priors} \end{figure}

Having weaker \So and \St priors did considerable widen the posterior
estimates of \St for all populations (Fig.~\ref{fig:s2-trends-priors}). The
Saint-Jean River, which had the least uncertain \St estimates among all
populations, had considerably larger uncertainties in \St while still
maintaining a very similar temporal trend in \St. The Trinité River also
maintained a similar temporal trend in \St, but the weaker priors resulted in
annual estimates of \St showing a larger spread, particularly in 1994 and 1995
which were the years withe highest estimated \St.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/s2-trends-priors.png}
    \caption{Median posterior estimates of survival in the second year at sea
        (\St) for the seven populations examined under three scenarios with
        different priors. The strong prior scenario is the one presented in
        our study, the weak \Pg priors uses the same weak prior for all
        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
    \label{fig:s2-trends-priors} \end{figure}

Estimates of \Pg under the weak \Pg scenario look almost identical to those
under the strong priors scenario, however the weak \So and \St scenario
resulted in more uncertain \Pg estimates for non 1SW-dominated populations
(Fig.~\ref{fig:pr-trends-priors}). For these populations, the temporal trends
observed in the strong prior scenario were maintained under the weak \So and
\St scenario, albeit weakened by the larger uncertainties in annual \Pg
estimates.

Overall, weakening the \prmu priors did not have an observable effect on model
parameter estimates, while weakening \So and \St did result in more uncertain
estimates of \St and \Pg, with estimates of \So being mostly unchanged.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/pr-trends-priors.png}
    \caption{Median posterior estimates of proportion returning after one winter at sea 
        (\Pg) for the seven populations examined under three scenarios with
        different priors. The strong prior scenario is the one presented in
        our study, the weak \Pg priors uses the same weak prior for all
        populations as shown by Fig.~\ref{fig:prweak}, while the weak \So \&
        \St scenario used wider survival priors as shown in Fig.~\ref{fig:zweak}.} 
    \label{fig:pr-trends-priors} \end{figure}

We calculated the correlations of estimated trends in $Z_1$ among populations
under the weak \So and \St scenario. While the direction of the correlation
stayed the same between all populations when compared with the strong prior
scenario, there were differences in the correlation coefficients (Fig.~\ref{fig:s1-corr-weakz}).
Among these, the correlation coefficients differed significantly for three pair-wise
comparisons: the correlations between Campbellton and Nashwaak as well
Campbellton and Trinité changed from non-significant to positively
significant, while the correlation between LaHave and Saint-Jean changed from negatively significant to
non-significant even though the coefficient value decreased further into the
negative.
Twelve out of the 21 pair-wise correlations were significant (p-value below 0.05) yet under the  
weak \So and \St scenario 9 out of the twelve significant correlations were positive.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.95\linewidth]{figures/corrplot-post-z1-corrected-weakz.png} \caption{
        Correlations of estimated trends in $Z_1$ among populations with weak \So and
\St priors. Correlation coefficients are shown in each square, while colouring
        denotes significance of the correlation ($p \leq 0.05$) as well as the direction and magnitude.}
\label{fig:s1-corr-weakz}
\end{figure}

%\clearpage
%\subsection*{References}

\nobibliography{subset}

\end{document}